---
title: "Example Code"
author: "S.C. McClelland"
date: "2026-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

Example code snippets for analysis scripts are available here. A full script example is available at the end of this document.

### Estimating change in soil organic carbon

```{r delta}
# data sign: (currently negative)
# units: Mg C and Mg CO2-eq
delta_soc_co2_eq          = function(delta_somsc) { # negative output indicates soil C sink
  gmwCO2 = 44
  gmwC   = 12
  Mg_ha  = 100
  delta_soc_co2_eq        = ((-delta_somsc)*(gmwCO2/gmwC))/Mg_ha 
}
# convert from Mg C to Mg CO2-eq
dt[, s_SOC   := delta_soc_co2_eq(s_somsc)]

#------------------------------------------------------------------------------
# ESTIMATE DELTA (always load conv_dt)
#------------------------------------------------------------------------------
# delta is scenario - conv
{
  ## example for res ##
d_res_dt = res_dt[conv_dt[, -c('scenario')], on = .(gridid   = gridid,
                                                            crop     = crop,
                                                            irr      = irr,
                                                            ssp      = ssp,
                                                            gcm      = gcm,
                                                            y_block  = y_block,
                                                            rep      = rep)]
rm(res_dt, conv_dt) # save space
gc()
# compute d_s_SOC (delta SOC)
d_res_dt[, d_s_SOC := s_SOC - i.s_SOC]
# remove absolute values (we only want difference)
d_res_dt[, i.s_SOC := NULL] 
d_res_dt[, s_SOC := NULL]
}
```

### Statistics

```{r monte carlo}
# compute mean, median, quantile / percentile responses for each iteration / rep
# mean
dt_m = dt[, lapply(.SD, mean), .SDcols = 'd_s_SOC',
                    by = .(rep)]
# 25th percentile
dt_q2 = dt[, lapply(.SD, function(x) {quantile(x)[2]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
```

### Full example

This is an example script for estimating statistics and visualizing data. It provides an example for global and two country-level estimates. We will also need to add in soil and environmental features for additional analyses (TBD).

The PDF and CDF approach will also need to be refined further.

```{r full example}
# filename:     histogram.R
# created:      22 March 2025
# last updated: 04 February 2026
# author:       S.C. McClelland

# description:  This script computes statistics, pdfs, and cdfs. 
#----------------------------------------------------------------
# libraries
#----------------------------------------------------------------
library(data.table)
library(ggplot2)
library(ggridges)
library(rstudioapi)
library(sf)
library(terra)
options(scipen = 999)
#----------------------------------------------------------------
# set base paths
#----------------------------------------------------------------
# get current directory, files should be in same directory
b_path = dirname(getActiveDocumentContext()$path)
d_path = paste(b_path, 'data', sep = '/')
shp_p  = paste(b_path, 'shp', sep = '/')
#----------------------------------------------------------------
# load data tables
#----------------------------------------------------------------
# data overview:
# scenario   residue retention
# y_block    2036 (21 years) 
# gridid     cell numbers
# crop       maiz (maize), soyb(soybean), swht (spring wheat), wwht (winter wheat)
# irr        0 (rainfed), 1 (irrigated)
# d_s_SOC    SOC (soil carbon) response, Mg CO2-eq ha-1
# Positive responses = higher emissions
# Negative responses = mitigation or emissions reductions
load(paste(d_path, 'res-delta-ghg-flux-uncertainty.RData', sep = '/'))
head(d_res_dt)
d_res_dt = d_res_dt[,c('gridid', 'crop', 'irr', 'rep', 'd_s_SOC')]
#----------------------------------------------------------------
# GLOBAL statistics
#----------------------------------------------------------------
# mean
res_mean = d_res_dt[, lapply(.SD, mean), .SDcols = 'd_s_SOC',
                    by = .(rep)]
# median
res_median = d_res_dt[, lapply(.SD, median), .SDcols = 'd_s_SOC',
                    by = .(rep)]
# quantile
res_q1 = d_res_dt[, lapply(.SD, function(x) {quantile(x)[1]}),
                  .SDcols = 'd_s_SOC',
                    by = .(rep)]
res_q2 = d_res_dt[, lapply(.SD, function(x) {quantile(x)[2]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
res_q3 = d_res_dt[, lapply(.SD, function(x) {quantile(x)[3]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
res_q4 = d_res_dt[, lapply(.SD, function(x) {quantile(x)[4]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
res_q5 = d_res_dt[, lapply(.SD, function(x) {quantile(x)[5]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
# combine statistics dt
res_stats = rbind(res_mean[,   stat := 'mean'],
                  res_median[, stat := 'median'],
                  res_q1[,     stat := 'min'],
                  res_q2[,     stat := '25th'],
                  res_q4[,     stat := '75th'],
                  res_q5[,     stat := 'max'])
#----------------------------------------------------------------
# ADD regions
#----------------------------------------------------------------
r_shp   = st_read(paste(shp_p, 'WB_countries_Admin0_10m.shp', sep = '/'))
r       = rast(paste(b_path, 'msw-cropland-rf-ir-area.tif', sep = '/'))
r       = r[[1]]
create_WB_cty = function(shp_f, rst) {
  # ORIGINAL WB NAME and OBJECTID
  shp_dt          = setDT(as.data.table(st_drop_geometry(shp_f)))
  # MATCH resolution of simulation data, dimensions the same
  target.r    = rst
  # CONFIRM SHP in same crs as target
  country.sf  = st_transform(shp_f, crs(target.r))
  country_r   = terra::rasterize(
    x         = vect(country.sf),
    y         = target.r,
    field     = "OBJECTID",
    touches   = TRUE          # optional: include cells touched by polygons
  )
  # CREATE data.frame, merge
  new_shp_dt  = as.data.frame(country_r, cells = TRUE, xy = TRUE)
  new_shp_dt  = setDT(new_shp_dt) # data.table object
  # GET WB names to match to ID
  shp_names   = data.table(WB_NAME   = shp_dt$WB_NAME,
                           ID        = shp_dt$OBJECTID)
  # JOIN with WB names
  new_shp_dt  = new_shp_dt[shp_names, on = .(OBJECTID = ID)]  
  return(new_shp_dt)
}
WB_dt = create_WB_cty(r_shp, r)
# join
d_res_dt = d_res_dt[WB_dt[,c('cell', 'WB_NAME')], on = .(gridid = cell)]
d_res_dt = d_res_dt[!is.na(crop)]
setorder(d_res_dt, gridid)
gc()
#----------------------------------------------------------------
# US statistics
#----------------------------------------------------------------
US_d_res_dt = d_res_dt[WB_NAME == 'United States of America']
# mean
US_res_mean = US_d_res_dt[, lapply(.SD, mean), .SDcols = 'd_s_SOC',
                    by = .(rep)]
# median
US_res_median = US_d_res_dt[, lapply(.SD, median), .SDcols = 'd_s_SOC',
                      by = .(rep)]
# quantile
US_res_q1     = US_d_res_dt[, lapply(.SD, function(x) {quantile(x)[1]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
US_res_q2     = US_d_res_dt[, lapply(.SD, function(x) {quantile(x)[2]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
US_res_q4     = US_d_res_dt[, lapply(.SD, function(x) {quantile(x)[4]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
US_res_q5     = US_d_res_dt[, lapply(.SD, function(x) {quantile(x)[5]}),
                  .SDcols = 'd_s_SOC',
                  by = .(rep)]
# combine statistics dt
US_res_stats = rbind(US_res_mean[,   stat := 'mean'],
                     US_res_median[, stat := 'median'],
                     US_res_q1[,     stat := 'min'],
                     US_res_q2[,     stat := '25th'],
                     US_res_q4[,     stat := '75th'],
                     US_res_q5[,     stat := 'max'])
#----------------------------------------------------------------
# Brazil statistics
#----------------------------------------------------------------
BR_d_res_dt = d_res_dt[WB_NAME == 'Brazil']
# mean
BR_res_mean = BR_d_res_dt[, lapply(.SD, mean), .SDcols = 'd_s_SOC',
                          by = .(rep)]
# median
BR_res_median = BR_d_res_dt[, lapply(.SD, median), .SDcols = 'd_s_SOC',
                            by = .(rep)]
# quantile
BR_res_q1     = BR_d_res_dt[, lapply(.SD, function(x) {quantile(x)[1]}),
                            .SDcols = 'd_s_SOC',
                            by = .(rep)]
BR_res_q2     = BR_d_res_dt[, lapply(.SD, function(x) {quantile(x)[2]}),
                            .SDcols = 'd_s_SOC',
                            by = .(rep)]
BR_res_q4     = BR_d_res_dt[, lapply(.SD, function(x) {quantile(x)[4]}),
                            .SDcols = 'd_s_SOC',
                            by = .(rep)]
BR_res_q5     = BR_d_res_dt[, lapply(.SD, function(x) {quantile(x)[5]}),
                            .SDcols = 'd_s_SOC',
                            by = .(rep)]
# combine statistics dt
BR_res_stats = rbind(BR_res_mean[,   stat := 'mean'],
                     BR_res_median[, stat := 'median'],
                     BR_res_q1[,     stat := 'min'],
                     BR_res_q2[,     stat := '25th'],
                     BR_res_q4[,     stat := '75th'],
                     BR_res_q5[,     stat := 'max'])
#----------------------------------------------------------------
# create ridgeline plots
#----------------------------------------------------------------
# GLOBAL, ALL CROPS, IRR
# order
res_stats$stat = factor(res_stats$stat, levels = c('min', '25th', 'mean', 'median','75th', 'max'))
stat_label     = c('Minimum', '25th Percentile', 'Mean', 'Median', '75th Percentile', 'Maximum')

# annual values
t_period     = 21L
CO2e_C       = (44/12)
s_flip       = -1L
# Violin plot
# Combine scale + bandwidth
# scale: reduce the height and overlap of the ridges
# bandwidth: Higher bandwidth = smoother, lower peaks
ggplot(res_stats, aes(x = (((d_s_SOC)/t_period)/CO2e_C)*s_flip, y = stat, fill = stat)) +
  geom_density_ridges(alpha = 0.7, scale = 0.6, rel_min_height = 0.005,
                      bandwidth = 0.03) +
  geom_vline(xintercept = 0, linetype = "solid", color = "black") +
  scale_y_discrete(expand = c(0.01, 0), 
                   labels = stat_label) +
  scale_x_continuous(breaks = seq(-0.25, 3, by = 0.25)) +  # Every 0.25
  labs(title = "Uncertainty in Soil Carbon Change Statistics, 20-years",
       subtitle = "Distribution across 1001 Monte Carlo iterations",
       x = "Global Soil Carbon Change (Mg C ha-1 yr-1)",
       y = "Summary Statistic") +
  theme_bw() +
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14),
        axis.text   = element_text(size = 12),
        plot.title = element_text(size = 18),
        plot.subtitle = element_text(size = 16))
# US, ALL CROPS, IRR
# order
US_res_stats$stat = factor(US_res_stats$stat, levels = c('min', '25th', 'mean', 'median','75th', 'max'))
stat_label     = c('Minimum', '25th Percentile', 'Mean', 'Median', '75th Percentile', 'Maximum')

# annual values
t_period     = 21L
CO2e_C       = (44/12)
# Violin plot
# Combine scale + bandwidth
# scale: reduce the height and overlap of the ridges
# bandwidth: Higher bandwidth = smoother, lower peaks
ggplot(US_res_stats, aes(x = (((d_s_SOC)/t_period)/CO2e_C)*s_flip, y = stat, fill = stat)) +
  geom_density_ridges(alpha = 0.7, scale = 0.6, rel_min_height = 0.005,
                      bandwidth = 0.03) +
  geom_vline(xintercept = 0, linetype = "solid", color = "black") +
  scale_y_discrete(expand = c(0.01, 0), 
                   labels = stat_label) +
  scale_x_continuous(breaks = seq(-0.1, 1, by = 0.1)) +  # Every 0.1
  labs(title = "Uncertainty in Soil Carbon Change Statistics, 20-years",
       subtitle = "Distribution across 1001 Monte Carlo iterations",
       x = "U.S. Soil Carbon Change (Mg C ha-1 yr-1)",
       y = "Summary Statistic") +
  theme_bw() +
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14),
        axis.text   = element_text(size = 12),
        plot.title = element_text(size = 18),
        plot.subtitle = element_text(size = 16))

# BRAZIL, ALL CROPS, IRR
# order
BR_res_stats$stat = factor(BR_res_stats$stat, levels = c('min', '25th', 'mean', 'median','75th', 'max'))
stat_label     = c('Minimum', '25th Percentile', 'Mean', 'Median', '75th Percentile', 'Maximum')

# annual values
t_period     = 21L
CO2e_C       = (44/12)
# Violin plot
# Combine scale + bandwidth
# scale: reduce the height and overlap of the ridges
# bandwidth: Higher bandwidth = smoother, lower peaks
ggplot(BR_res_stats, aes(x = (((d_s_SOC)/t_period)/CO2e_C)*s_flip, y = stat, fill = stat)) +
  geom_density_ridges(alpha = 0.7, scale = 0.6, rel_min_height = 0.005,
                      bandwidth = 0.03) +
  geom_vline(xintercept = 0, linetype = "solid", color = "black") +
  scale_y_discrete(expand = c(0.01, 0), 
                   labels = stat_label) +
  scale_x_continuous(breaks = seq(-0.25, 3, by = 0.25)) +  # Every 0.25
  labs(title = "Uncertainty in Soil Carbon Change Statistics, 20-years",
       subtitle = "Distribution across 1001 Monte Carlo iterations",
       x = "Brazil Soil Carbon Change (Mg C ha-1 yr-1)",
       y = "Summary Statistic") +
  theme_bw() +
  theme(legend.position = "none",
        axis.ticks.y = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_text(size = 14),
        axis.text   = element_text(size = 12),
        plot.title = element_text(size = 18),
        plot.subtitle = element_text(size = 16))
#----------------------------------------------------------------
# CDF / PDF
#----------------------------------------------------------------
# # Pool all data across iterations for the full distribution
# all_data = d_res_dt[, .(d_s_SOC)]
# 
# # CDF plot
# ggplot(all_data, aes(x = ((d_s_SOC)/t_period)/CO2e_C)) +
#   stat_ecdf(geom = "step", linewidth = 1, color = "steelblue") +
#   geom_vline(xintercept = 0, linetype = "dashed", color = "red", linewidth = 0.8) +
#   geom_hline(yintercept = c(0.05, 0.5, 0.95), 
#              linetype = "dotted", color = "gray50", alpha = 0.6) +
#   annotate("text", x = min(all_data$d_s_SOC) * 0.9, y = 0.05, 
#            label = "5th percentile", hjust = 0, size = 3) +
#   annotate("text", x = min(all_data$d_s_SOC) * 0.9, y = 0.5, 
#            label = "Median", hjust = 0, size = 3) +
#   annotate("text", x = min(all_data$d_s_SOC) * 0.9, y = 0.95, 
#            label = "95th percentile", hjust = 0, size = 3) +
#   scale_y_continuous(labels = scales::percent_format()) +
#   labs(title = "Cumulative Distribution of Soil Carbon Change",
#        subtitle = "Pooled across all Monte Carlo iterations",
#        x = "Soil Carbon Change (Mg C/ha)",
#        y = "Cumulative Probability") +
#   theme_minimal() +
#   theme(panel.grid.minor = element_blank())
# 
# 
# # PDF with shaded tail regions (extremes)
# 
# # Calculate percentiles
# p05 = quantile(all_data$d_s_SOC, 0.05)
# p95 = quantile(all_data$d_s_SOC, 0.95)
# 
# # Get density for shading
# dens = density(all_data$d_s_SOC)
# dens_dt = data.table(x = dens$x, y = dens$y)
# 
# ggplot(all_data, aes(x = ((d_s_SOC)/t_period)/CO2e_C)) +
#   # Shade lower 5% tail
#   geom_area(data = dens_dt[x <= p05], 
#             aes(x = x, y = y), fill = "red", alpha = 0.4) +
#   # Shade upper 5% tail
#   geom_area(data = dens_dt[x >= p95], 
#             aes(x = x, y = y), fill = "darkgreen", alpha = 0.4) +
#   geom_density(fill = "steelblue", alpha = 0.3, linewidth = 1) +
#   geom_vline(xintercept = 0, linetype = "dashed", color = "red") +
#   labs(title = "Probability Density of Soil Carbon Change",
#        subtitle = "Shaded regions = worst 5% (red) and best 5% (green)",
#        x = "Soil Carbon Change (Mg C/ha)",
#        y = "Density") +
#   theme_minimal()
```
